\normalfont\textbf{ \Large {Reducing data traffic in wireless structural health monitoring systems using embedded computing}}

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
abstract

@@

\renewcommand\refname{References and Notes}

@@

\title{\textbf{ \Large{Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}

\author
{Katrin Jahr, Robert Schlich\\
\\
\normalsize{Degree program “Civil Engineering” (M.Sc.)}\\
\normalsize{Bauhaus-Universität Weimar, Germany}\\
\\
\normalsize{katrin.jahr@uni-weimar.de}\\
\normalsize{robert.schlich@uni-weimar.de}
}

@@

\baselineskip18pt

@@

The dependability and accuracy of wireless structural health monitoring systems can be affected by data loss during wireless data transmission.
Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and lower the occurrence of data loss.
In this paper, the design and implementation of a wireless structural health monitoring system, capable of data compression is presented. 
A fast Fourier transform is used on every sensor node to analyze oscillation frequencies. 
Only the first natural frequency of the oscillation and the correlated magnitude are transmitted to the base station.
In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced by xx %.

@@

\linenumbers % Schaltet Zeilennummerierung ein
\modulolinenumbers[5] % nur jede 5. Zeile

\section*{Dictionary}

\begin{tabular}{|l|l|}
\hline 
Sensorknoten (SunSPOT) & sensor node \\ 
\hline 
einzelner Messsensor (Thermometer) & sensor \\ 
\hline 
Knoten im neuronalen Netz & neuron \\ 
\hline 
eine abgeschlossene Messungreihe & test run \\ 
\hline 
gemessene Werte & sensor data \\ 
\hline 
vorhergesagte Werte & predicted data \\ 
\hline 
durch Vorhersage erwartete Werte & expected data \\ 
\hline 
einzelner Messwert & measurement \\ 
\hline 
Test & laboratory experiments\\ 
\hline
tatsächliche, nicht virtuelle Messung & actual measurement \\
\hline
Messaufbau & test setup \\
\hline
\end{tabular}

@@

However, when using wireless SHM systems, special attention has to be paid to data transfer.
Immense amounts of raw data incure and have to be transmitted.
Over the air transmission is error-prone and may lead to irreparable data loss.  (klingt das zu stark? Warum ist das mit FFT nicht mehr so schlimm?)
To save transmission time and storage space, and to decrease data loss, the transmidded data can be reduced using embedded computing.
Depending on the measuring objective, first steps of data processing can be conducted decentralized directly on the sensor nodes.
Transmitting analysis results instead of raw data can reduce the amount of transmitted data exponentially.

@@

\newpage

>>>>

\makeatletter
\renewcommand\section{\@startsection{section}{1}{\z@}%
                     {-3.5ex \@plus -1ex \@minus -.2ex}%
                     {2.3ex \@plus.2ex}%
                     {\normalfont\large\bfseries}}
\makeatother

%\usepackage{titlesec}
%\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}

@@

%\renewcommand\refname{References and Notes}

@@

\title{\normalfont\textbf{ \Large {Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}

\author
{\large Katrin Jahr, Robert Schlich\\
\\
\normalsize{Degree program “Civil Engineering” (M.Sc.)}\\
\normalsize{Bauhaus-Universität Weimar, Germany}\\
\\
\normalsize{katrin.jahr@uni-weimar.de, }
\normalsize{robert.schlich@uni-weimar.de}
}

@@

\baselineskip20pt

@@

Immense amounts of data incure during wireless structural health monitoring.
Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and stored.
In this paper, the design and implementation of a wireless structural health monitoring system, capable of dezentralized data compression, is presented. 
A fast Fourier transform is embedded into each sensor node to analyze oscillation frequencies. 
By only transmitting the first natural frequency and the correlating magnitude, the data stream is reduced to two values.
The proposed approach was validated in laboratory experiments.

%The dependability and the accuracy of wireless structural health monitoring systems can be affected by data loss during wireless data transmission.
%Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and lower the occurrence of data loss.
%In this paper, the design and implementation of a wireless structural health monitoring system, capable of data compression, is presented. 
%A fast Fourier transform is used on every sensor node to analyze oscillation frequencies. 
%Only the first natural frequency of the oscillation and the correlated magnitude are transmitted to the base station.
%In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced by xx.

@@

\linenumbers % Schaltet Zeilennummerierung ein
\modulolinenumbers[5] % nur jede 5. Zeile
%
%\section*{Dictionary}
%
%\begin{tabular}{|l|l|}
%\hline 
%Sensorknoten (SunSPOT) & sensor node \\ 
%\hline 
%einzelner Messsensor (Thermometer) & sensor \\ 
%\hline 
%Knoten im neuronalen Netz & neuron \\ 
%\hline 
%eine abgeschlossene Messungreihe & test run \\ 
%\hline 
%gemessene Werte & sensor data \\ 
%\hline 
%vorhergesagte Werte & predicted data \\ 
%\hline 
%durch Vorhersage erwartete Werte & expected data \\ 
%\hline 
%einzelner Messwert & measurement \\ 
%\hline 
%Test & laboratory experiments\\ 
%\hline
%tatsächliche, nicht virtuelle Messung & actual measurement \\
%\hline
%Messaufbau & test setup \\
%\hline
%\end{tabular}

@@

However, when using wireless SHM systems, special attention has to be paid to data transfer.
Immense amounts of raw data incure and have to be transmitted.
Over the air transmission is error-prone and may lead to irreparable data loss.  (klingt das zu stark? Warum ist das mit FFT nicht mehr so schlimm?)
To save transmission time and storage space, and to decrease data loss, the transmitted data can be reduced using embedded computing.
Depending on the measuring objective, first steps of data processing can be conducted decentralized directly on the sensor nodes.
Transmitting analysis results instead of raw data can reduce the amount of transmitted data exponentially.

@@

%\newpage
